---
title: "Lab 05 - Wrangling spatial data"
author: "Allison Li"
date: "02/09/2025"
output: 
  html_document: 
    toc: yes
    toc_float: yes
---

```{r load-pkg, message = FALSE}
##devtools::install_github("rstudio-education/dsbox")
library(tidyverse)
library(dsbox)
```

### Exercise 1
```{r filtered AK data}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)

lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(dn_ak)
```

There are `r nrow(dn_ak)` Denny's stores in Alaska and `r nrow(lq_ak)` La Quinta hotels in Alaska.

### Exercise 2

```{r}
pairings_of_lqak <- nrow(dn_ak) * nrow(lq_ak)
cat("The number of pairings between the locations of Denny’s and La Quinta hotels in Alaska are:", pairings_of_lqak, "\n")
```


### Exercise 3

```{r load-data, message = FALSE}
# From TidyTuesday: https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-02-11/readme.md
hotels <- read.csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-11/hotels.csv")

```

First, knit the document and view the following visualization. How are the months ordered? What would be a better order? Then, reorder the months on the x-axis (levels of `arrival_date_month`) in a way that makes more sense. You will want to use a function from the **forcats** package, see https://forcats.tidyverse.org/reference/index.html for inspiration and help.

**Stretch goal:** If you finish the above task before time runs out, change the y-axis label so the values are shown with dollar signs, e.g., $80 instead of 80. You will want to use a function from the **scales** package. For inspiration and help, see https://scales.r-lib.org/reference/index.html.

```{r plot, fig.width=10}
hotels %>%
  group_by(hotel, arrival_date_month) %>%   # group by hotel type and arrival month
  summarise(mean_adr = mean(adr)) %>%       # calculate mean adr for each group
  ggplot(aes(
    x = arrival_date_month,                 # x-axis = arrival_date_month
    y = mean_adr,                           # y-axis = mean_adr calculated above
    group = hotel,                          # group lines by hotel type
    color = hotel)                          # and color by hotel type
    ) +
  geom_line() +                             # use lines to represent data
  theme_minimal() +                         # use a minimal theme
  labs(x = "Arrival month",                 # customize labels
       y = "Mean ADR (average daily rate)",
       title = "Comparison of resort and city hotel prices across months",
       subtitle = "Resort hotel prices soar in the summer while ciry hotel prices remain relatively constant throughout the year",
       color = "Hotel type")
```

## References {#references}
1. Assignment Adapted from [Mine Cetinkaya-Rundel's Data Science in a Box](https://github.com/rstudio-education/datascience-box)
