---
title: "Lab 05 - Wrangling spatial data"
author: "Allison Li"
date: "02/09/2025"
output: 
  html_document: 
    toc: yes
    toc_float: yes
---

```{r load-pkg, message = FALSE}
##devtools::install_github("rstudio-education/dsbox")
library(tidyverse)
library(dsbox)
```

### Exercise 1
```{r filtered AK data}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)

lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(dn_ak)
```

There are `r nrow(dn_ak)` Denny's stores in Alaska and `r nrow(lq_ak)` La Quinta hotels in Alaska.

### Exercise 2

```{r}
pairings_of_lqak <- nrow(dn_ak) * nrow(lq_ak)
cat("The number of pairings between the locations of Denny’s and La Quinta hotels in Alaska are:", pairings_of_lqak, "\n")
```
I asked ChatGPT to help me generate this code because I was not sure what could be the correct function for this.

### Exercise 3

```{r load-data, message = FALSE}

dn_lq_ak <- full_join(dn_ak, lq_ak, 
                      by = "state")
dn_lq_ak
```

### Exercise 4

There are `r ncol(dn_lq_ak)` observations in the joined data frame. The names including city.x, zip.x, longitude.x, latitude.x, address.x, which all refer to the locations of the Denny's. The names of the rest variables, including city.y, zip.y, longitude.y, latitude.y, address.y all refer to the locations of the La Quinta hotels. 

### Exercise 5
```{r haversine}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 <- long1 * pi / 180
  lat1 <- lat1 * pi / 180
  long2 <- long2 * pi / 180
  lat2 <- lat2 * pi / 180

  R <- 6371 # Earth mean radius in km

  a <- sin((lat2 - lat1) / 2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1) / 2)^2
  d <- R * 2 * asin(sqrt(a))

  return(round(d, round)) # distance in km
}
```

To answer the question of what function to use from the tidyverse to add a new variable while keeping the existing ones, I believe we can use mutate().

### Exercise 6
```{r distance between all pairs}

dn_lq_ak <- dn_lq_ak %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))

print(dn_lq_ak)
```

### Exercise 7

```{r minimum distance}

## I changed summarize to mutate here because I want closest distance to be a new variable in the dataset while keeping other varaibles as well.
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  mutate(closest = min(distance))

print(dn_lq_ak_mindist)
```

### Exercise 8

```{r report and visualization}
## Mean 
mean_closestdist <- mean(dn_lq_ak_mindist$closest) 
## Standard deviation 
sd_closestdist <- sd(dn_lq_ak_mindist$closest) 

## I want to merge the two dataset so that I can have a set including five stores and their location information, as well as which ones are La Quinta and which ones are Denny's.
dn_ak <- dn_ak %>% mutate(source = "Denny's")
lq_ak <- lq_ak %>% mutate(source = "La Quinta")
merged_data_ak <- bind_rows(dn_ak, lq_ak)

## The section block below is helped by ChatGPT. I asked it how to generate a dataset with only the three minimum distances and their longitude, latitude, and other variables of Denny's and La Quinta hotels accordingly.
dn_lq_ak_minidis <- dn_lq_ak %>%
  group_by(address.x) %>%
  filter(distance == min(distance)) %>% 
  ungroup()

## I want the visualized plot to be able to show the shortest distances between each Denny's and La Quinta hotel. 
ggplot(merged_data_ak, aes(x = longitude, y = latitude, color = source)) +
  geom_point(size = 3, alpha = 0.8) + 
  geom_segment(data = dn_lq_ak_minidis,
               aes(x = longitude.x, y = latitude.x, 
                   xend = longitude.y, yend = latitude.y), 
               color ="purple", linetype = "solid", size = .8)
  labs(title = "Scatter Plot of Denny's and La Quinta",
       x = "Longitude",
       y = "Latitude",
       color = "Dataset")

```

The three minimum distance between the Denny's and La Quinta hotels are `r round(sort(dn_lq_ak_mindist$closest)[1], 2)`, `r round(sort(dn_lq_ak_mindist$closest)[2], 2)`, and `r round(sort(dn_lq_ak_mindist$closest)[3], 2)` separately. The mean for the minimum distances is `r round(mean_closestdist, 2)` while the standard deviation is `r round(sd_closestdist, 2)`.

In terms of visiualization, with the help of Chatgpt, I am able to add the purple lines that demostrate the minimum distances between Denny's and La Quinta hotels, yay!

### Exercise 9.1
```{r filtered NC data}
dn_nc <- dennys %>%
  filter(state == "NC")
nrow(dn_nc)

lq_nc <- laquinta %>%
  filter(state == "NC")
nrow(dn_nc)
```
### Exercise 9.2
```{r all possible pairings}
pairings_of_lqnc <- nrow(dn_nc) * nrow(lq_nc)
cat("The number of pairings between the locations of Denny’s and La Quinta hotels in North Carolina are:", pairings_of_lqnc, "\n")
```
### Exercise 9.3
```{r all the distance between in NC}
dn_lq_nc <- full_join(dn_nc, lq_nc, 
                      by = "state")
dn_lq_nc

dn_lq_nc <- dn_lq_nc %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))

print(dn_lq_nc)
```
### Exercise 9.4
```{r minimum distances in NC}
dn_lq_nc_minidis <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

print(dn_lq_nc_minidis)
```
### Exercise 9.5
```{r}
## Mean 
mean_closestdist_nc <- mean(dn_lq_nc_minidis$closest) 
## Standard deviation 
sd_closestdist_nc <- sd(dn_lq_nc_minidis$closest) 

dn_nc <- dn_nc %>% mutate(source = "Denny's")
lq_nc <- lq_nc %>% mutate(source = "La Quinta")
merged_data_nc <- bind_rows(dn_nc, lq_nc)

## filter the closest distances
dn_lq_nc_minidis <- dn_lq_nc %>%
  group_by(address.x) %>%
  filter(distance == min(distance)) %>% 
  ungroup()

## A scattor plot as the visualization to show the shortest distances between each Denny's and La Quinta hotel. 
ggplot(merged_data_nc, aes(x = longitude, y = latitude, color = source)) +
  geom_point(size = 3, alpha = 0.8) + 
  geom_segment(data = dn_lq_nc_minidis,
               aes(x = longitude.x, y = latitude.x, 
                   xend = longitude.y, yend = latitude.y), 
               color ="#89CFF0", linetype = "solid", size = .8)
```

There are `r nrow(dn_lq_nc_minidis)` minimum distances between each Denny's and a La Quinta hotel. The mean for the minimum distances is `r round(mean_closestdist_nc, 2)` while the standard deviation is `r round(sd_closestdist_nc, 2)`.


### Exercise 10.1
```{r filtered TX data}
dn_tx <- dennys %>%
  filter(state == "TX")
nrow(dn_tx)

lq_tx <- laquinta %>%
  filter(state == "TX")
nrow(dn_tx)
```
### Exercise 10.2
```{r all possible pairings in TX}
pairings_tx <- nrow(dn_tx) * nrow(lq_tx)
cat("The number of pairings between the locations of Denny’s and La Quinta hotels in Texas are:", pairings_tx, "\n")
```
### Exercise 10.3
```{r all the distance between in TX}
dn_lq_tx <- full_join(dn_tx, lq_tx, 
                      by = "state")
dn_lq_tx

dn_lq_tx <- dn_lq_tx %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))

print(dn_lq_tx)
```
### Exercise 10.4
```{r minimum distances in TX}
dn_lq_tx_minidis <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

print(dn_lq_tx_minidis)
```
### Exercise 10.5
```{r visualization of the shorest distances between each Dennys and La Quinta}
## Mean 
mean_closestdist_tx <- mean(dn_lq_tx_minidis$closest) 
## Standard deviation 
sd_closestdist_tx <- sd(dn_lq_tx_minidis$closest) 

dn_tx <- dn_tx %>% mutate(source = "Denny's")
lq_tx <- lq_tx %>% mutate(source = "La Quinta")
merged_data_tx <- bind_rows(dn_tx, lq_tx)

## filter the closest distances
dn_lq_tx_minidis <- dn_lq_tx %>%
  group_by(address.x) %>%
  filter(distance == min(distance)) %>% 
  ungroup()

## A scattor plot as the visualization to show the shortest distances between each Denny's and La Quinta hotel. 
ggplot(merged_data_tx, aes(x = longitude, y = latitude, color = source)) +
  geom_point(size = 3, alpha = 0.8) + 
  geom_segment(data = dn_lq_tx_minidis,
               aes(x = longitude.x, y = latitude.x, 
                   xend = longitude.y, yend = latitude.y), 
               color = "blue", linetype = "solid", size = .8)
```

There are `r nrow(dn_lq_tx_minidis)` minimum distances between each Denny's and a La Quinta hotel. The mean for the minimum distances is `r round(mean_closestdist_tx, 2)` while the standard deviation is `r round(sd_closestdist_tx, 2)`.
Additionally, according to the visualized plot, in Texas there are a lot of La Quinta hotels that each Denny's has a La Quinta hotel near by. 

### Exercise 11.1-- Ohio

My undergraduate studies was in Ohio so I would like to know how La Quinta and Denny's are located in Ohio.

```{r filtered OH data}
dn_oh <- dennys %>%
  filter(state == "OH")
nrow(dn_oh)

lq_oh <- laquinta %>%
  filter(state == "OH")
nrow(dn_oh)
```
### Exercise 11.2
```{r all possible pairings in OH}
pairings_oh <- nrow(dn_oh) * nrow(lq_oh)
cat("The number of pairings between the locations of Denny’s and La Quinta hotels in Ohio are:", pairings_oh, "\n")
```
### Exercise 11.3
```{r all the distance between in OH}
dn_lq_oh <- full_join(dn_oh, lq_oh, 
                      by = "state")
dn_lq_oh

dn_lq_oh <- dn_lq_oh %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))

print(dn_lq_oh)
```
### Exercise 11.4
```{r minimum distances in OH}
dn_lq_oh_minidis <- dn_lq_oh %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

print(dn_lq_oh_minidis)
```
### Exercise 11.5
```{r visualization of the shorest distances between each Dennys and La Quinta in Ohio}
## Mean 
mean_closestdist_oh <- mean(dn_lq_oh_minidis$closest) 
## Standard deviation 
sd_closestdist_oh <- sd(dn_lq_oh_minidis$closest) 

dn_oh <- dn_oh %>% mutate(source = "Denny's")
lq_oh <- lq_oh %>% mutate(source = "La Quinta")
merged_data_oh <- bind_rows(dn_oh, lq_oh)

## filter the closest distances
dn_lq_oh_minidis <- dn_lq_oh %>%
  group_by(address.x) %>%
  filter(distance == min(distance)) %>% 
  ungroup()

## A scattor plot as the visualization to show the shortest distances between each Denny's and La Quinta hotel. 
ggplot(merged_data_oh, aes(x = longitude, y = latitude, color = source)) +
  geom_point(size = 3, alpha = 0.8) + 
  geom_segment(data = dn_lq_oh_minidis,
               aes(x = longitude.x, y = latitude.x, 
                   xend = longitude.y, yend = latitude.y), 
               color = "#FFEE8C", linetype = "solid", size = .8)
```

There are `r nrow(dn_lq_oh_minidis)` minimum distances between each Denny's and a La Quinta hotel. The mean for the minimum distances is `r round(mean_closestdist_oh, 2)` while the standard deviation is `r round(sd_closestdist_oh, 2)`.
Additionally, according to the visualized plot, there are alot of Denny's in Ohio and many of them do not have a La Quinta hotel near by. 

### Exercise 12

Amomg all the states I have examined, I believe Mitch Hedberg’s joke that each Denny's has a La Quinta nearby is most likely to hold true in Texas and Alaska. According to the data, Texas' mean value for closest distance between Denny's and La Quinta is 5.79, while the mean value in Alaska is 4.41. Based on the scattor plot, we can see that Denny's in both Texas and Alaska have at least one really close La Quinta hotel. 
